---
title: "high-dimentional time series project"
author: "Jesse van der Ende, Amine Ziddi and Lena Monsch"
date: "2026-01-29"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## read the datasets

# first option to work with:

- hourly PM2.5 measurements taken in March 2017 in Taiwan (in micrograms per cubic meters, μg/m2)
- 516 series each with 31 x 24 = 744 observations (i.e. k= 516 and T = 744)
```{r}
loc <- read.csv("locations032017.csv")
air <- read.csv("TaiwanAirBox032017.csv", col.names = c("time", paste0("col",2:517)))
head(loc)
head(air[,1:5])

```

```{r, echo = TRUE}
plot(air[,2], type ="l", ylab = "PM 2.5", ylim = c(min(air[-1]), max(air[,-1])), 
     col = rgb(0,0,0,alpha =0.05))
apply(air[,3:ncol(air)],2, function(i) points(i, type = "l", col = rgb(0,0,0,alpha =0.05)))

```

# second option to work with:
- weekly series of electricity price each hour of each day during 678 weeks in eight regions of New England (7 × 24 × 8 = 1344)
- Note that here k > T

L: but how are we supposed to know which data belongs to which region

```{r}
elec <- read.csv("PElectricity1344.csv", col.names = paste0("col",1:1344))
head(elec[,1:5])
```
```{r, echo = TRUE}
plot(elec[,1], type ="l", ylab = "electricity price", ylim = c(min(elec), max(elec)), 
     col = rgb(0,0,0,alpha =0.05))
apply(elec[2:ncol(elec)],2, function(i) points(i, type = "l", col = rgb(0,0,0,alpha =0.05)))


```

